---
title: 'CMTH 642 Data Analytics: Advanced Methods Assignment 1'
author: Paul Ycay
output:
  pdf_document: default
  word_document: default
---
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
### 1. Read the csv files in the folder. (4 points)
```{r}
Macro_Data<-read.csv(file="C:\\Users\\Paul\\Desktop\\USDA_Macronutrients.csv",header=T,sep=",")
Micro_Data<-read.csv(file="C:\\Users\\Paul\\Desktop\\USDA_Micronutrients.csv",header=T,sep=",")
head(Macro_Data)
head(Micro_Data)
```

### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (4 points)
```{r}
USDA<-merge(Macro_Data,Micro_Data,by="ID")
head(USDA)
```

### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (6 points)
#{r, eval=F, echo=T} #use this piece of code to only run the code  and not output
```{r}
sapply(USDA,class) 
USDA$Sodium<-gsub(',','',USDA$Sodium)
USDA$Potassium<-gsub(',','',USDA$Potassium)
USDA$Sodium<-as.numeric(USDA$Sodium)
USDA$Potassium<-as.numeric(USDA$Potassium)
sapply(USDA,class)

```

### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (6 points)
```{r}
missingvalues=(rowSums(is.na(USDA)))
USDA=USDA[!missingvalues > 4,];
sprintf("These are the number of records remaining: %i ",nrow(USDA))


```

### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (6 points)
```{r}
USDA$Sugar[is.na((USDA$Sugar))]<-mean(USDA$Sugar,na.rm = TRUE)
USDA$VitaminE[is.na((USDA$VitaminE))]<-mean(USDA$VitaminE,na.rm = TRUE)
USDA$VitaminD[is.na((USDA$VitaminD))]<-mean(USDA$VitaminD,na.rm = TRUE)


head(USDA)

```

### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (6 points)
```{r}
USDAclean=USDA[complete.cases(USDA),]
str(USDAclean)
sprintf("Number of records remaining: %i", nrow(USDAclean))
```

### 7. Which food has the highest sodium level? (6 points)
```{r}
which.max(USDAclean$Sodium)
USDAclean$Description[265]
```

### 8. Create a histogram of Vitamin C distribution in foods, with a limit of 0 to 100 on the x-axis and breaks of 100. (6 points)
```{r}
hist(USDAclean$VitaminC, xlab="Vitamin c", ylab="Milligrams (mg)", 
main= "Vitamin C Distribution", xlim=c(0,100), breaks=100)
```

### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (6 points)
```{r}
boxplot(USDAclean$Carbohydrate, USDAclean$TotalFat, USDAclean$Protein, 
main="Boxplot of Total Fat, 
Protein, &
Carbs", ylab="Grams (g)", 
xlab=("Carbohydrates             Total Fat              Protein"))
```

### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its calorie content. (6 points)
```{r}
plot(USDAclean$TotalFat, USDAclean$Calories, 
xlab="Total Fat", ylab = "Calories", 
main = "Total Fat vs Calories", col = "yellow")

```

### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise.Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (8 points)
```{r}
HighSodium = as.numeric(USDAclean$Sodium > mean(USDAclean$Sodium, na.rm=TRUE))
str(HighSodium)

HighCalories=as.numeric(USDAclean$Calories > mean(USDAclean$Calories,na.rm=TRUE))
str(HighCalories)  

HighProtein = as.numeric(USDAclean$Protein > mean(USDAclean$Protein,na.rm=TRUE))
str(HighProtein)

HighSugar = as.numeric(USDAclean$Sugar > mean(USDAclean$Sugar, na.rm=TRUE))
str(HighSugar)

HighFat = as.numeric(USDAclean$TotalFat > mean(USDAclean$TotalFat, na.rm=TRUE))

a<-table(HighSodium, HighFat);
highfs<-a[2,2];
paste0("Number of foods with high sodium and high fat: ", highfs)

```

### 12. Calculate the average amount of iron, sorted by high and low protein. (8 points)
```{r}
tapply(USDAclean$Iron, HighProtein, mean, na.rm=TRUE)
```

### 13. Create a script for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis for this script. (8 points)
```{r}
require(jpeg)
img<-readJPEG("C:\\Users\\Paul\\Desktop\\HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
healthcheck<- function(x,y,z)
{ifelse(x==1,ifelse(y==1,ifelse(z==1,"Fail","Pass"),"Pass"),"Pass")}
```

### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (8 points)
```{r}
USDAclean["HealthCheck"]<-healthcheck(HighSodium,HighSugar,HighFat)
head(USDAclean)
tail(USDAclean)
```

### 15. How many foods in the USDAclean data frame fail the HealthCheck? (8 points)
```{r}
nasty_foods<-sum(USDAclean$HealthCheck=="Fail", na.rm = TRUE)
paste0("Number of foods that fail the HealthCheck: ",nasty_foods)
```

### 16. Save your final data frame as "USDAclean_ [your last name]" (4 points)
```{r}
USDAclean_Ycay<-USDAclean
head(USDAclean_Ycay)
tail(USDAclean_Ycay)
```


