---
title: "CMTH642 Assignment 3"
author: "Paul Ycay"
date: "25/07/2019"
output:
  word_document: default
  pdf_document: default
course: CMTH 642
---
The following document was used to supplement this project https://rpubs.com/shradhit/winequality
The RMD file for Lab 10 and Lab 10 solutions was used to supplement this assignment as well
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
install.packages("ggplot2",
   repos = c("http://rstudio.org/_packages",
   "http://cran.rstudio.com"))
install.packages("class",
   repos = c("http://rstudio.org/_packages",
   "http://cran.rstudio.com"))
install.packages("gmodels",
   repos = c("http://rstudio.org/_packages",
   "http://cran.rstudio.com"))
library(corrplot)
library(class)
library(gmodels)

```
1. Check data characteristics. Is there missing data? 

```{r, tidy=TRUE} 
wine<-read.csv(file="http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", header = TRUE, sep= ";");
str(wine)
head(wine)
tail(wine)
summary(wine)

sum(is.na(wine))
```
There is no missing data

2.What is the	correlation	between	the	attributes other than wine quality?

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
 
    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
 
    text(0.5, 0.5, txt, cex = cex * r) 
    text(.8, .8, Signif, cex=cex, col=2) 
}
wine_cor<-subset(wine, select=c(1:11))
pairs(wine_cor, lower.panel=panel.smooth, upper.panel=panel.cor)
corrplot(cor(wine_cor))
```

3. Graph the frequency distribution	of wine quality.
```{r}
hist(wine$quality)
```

4. Reduce the levels of rating for quality to three levels as high, medium and	
low.


```{r}
wine$quality = ifelse(wine$quality < 5, 'low', ifelse(wine$quality > 7, 'high', 'medium'))
wine$quality = ordered(wine$quality, c('low', 'medium', 'high'))
round(prop.table(table(wine$quality)) * 100, digits = 1)
head(wine$quality)
tail(wine$quality)
summary(wine$quality)
```
5. Normalize the data set. 
```{r}
normalize <- function(x) {
               return ((x - min(x)) / (max(x) - min(x))) }
wine_n <- as.data.frame(lapply(wine[-12],normalize))
wine_n <- cbind(wine_n,wine$quality)
head(wine_n)
tail(wine_n)
summary(wine_n)
```
6. Divide the data to training and testing groups. 
```{r}
set.seed(1)
index <- sample(1:nrow(wine_n), 0.65 *nrow(wine_n))
wine_train <- wine_n[index,]
wine_test <- wine_n[-index,]
wine_train_labels <- wine_train[,12]
wine_test_labels <- wine_test[,12]
summary(wine_train_labels)
summary(wine_test_labels)
```
7. Use the KNN algorithm to predict quality of wine using its attributes 
```{r}
wine_test_pred <- knn(train = wine_train[,1:11], test = wine_test[,1:11],cl = wine_train[,1], k=10)
head(wine_test_pred)
tail(wine_test_pred)
summary(wine_test_pred)
```
8. Evaluate the model performance  
```{r}
CrossTable(x=wine_test_labels, y=wine_test_pred, prop.chisq=FALSE)
```
There were 1715 observations


